### 2022-04-23 README.qmd
Rename constraint_p to constraint_production #12
Former-commit-id: 4529e7d21166bac415501a34da0050828d3dfdb6

---
# format: html
format: gfm
bibliography: vignettes/references.bib
execute:
  cache: false
  fig.path: "man/figures/README-"
  message: false
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache = TRUE,
  fig.path = "man/figures/README-",
  message = FALSE)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/robinlovelace/si/workflows/R-CMD-check/badge.svg)](https://github.com/robinlovelace/si/actions)
<!-- badges: end -->

The goal of {si} is to provide a simple, '[tidy](https://dplyr.tidyverse.org/)', and flexible framework for developing spatial interaction models (SIMs), which estimate the amount of movement between spatial entities.

## Installation

Install the package as follows:

```r
install.packages("remotes") # if not already installed
```

```r
remotes::install_github("robinlovelace/si")
```

<!-- # Implementations in other languages -->
## si basics

Run a basic SIM as follows:

```{r distance}
library(si)
library(dplyr)
# prepare OD data
od = si_to_od(
  origins = si_zones,        # origin locations
  destinations = si_zones,   # destination locations
  max_dist = 5000            # maximum distance between OD pairs
  )
# specify a function
gravity_model = function(beta, d, m, n) {
  m * n * exp(-beta * d / 1000)
} 
# perform SIM
od_res = si_calculate(
  od,
  fun = gravity_model,
  d = distance_euclidean,
  m = origin_all,
  n = destination_all,
  constraint_production = origin_all,
  beta = 0.3
  )
# visualize the results
plot(od_res$distance_euclidean, od_res$interaction)
```

What just happened?
We created an 'OD data frame' with the function `si_to_od()` from geographic origins and destinations, and then estimated a simple 'production constrained' (with the `constraint_production` argument) gravity model based on the population in origin and destination zones and a custom distance decay function with `si_calculate()`.
As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.

The approach is also 'tidy', allowing use of {si} functions in {dplyr} pipelines:

```{r}
od_res = od %>% 
  si_calculate(fun = gravity_model, 
               m = origin_all,
               n = destination_all,
               d = distance_euclidean,
               constraint_production = origin_all,
               beta = 0.3)
od_res %>% 
  select(interaction)
```

The resulting estimates of interaction, returned in the column `interaction` and plotted with distance in the graphic above, resulted from our choice of spatial interaction model inputs, allowing a wide range of alternative approaches to be implemented.
This flexibility is a key aspect of the package, enabling small and easily modified functions to be implemented and tested.

The output `si_calculate)` is a geographic object which can be plotted as a map:

```{r map}
plot(od_res["interaction"], logz = TRUE)
```

The `si_to_od()` function transforms geographic entities (typically polygons and points) into a data frame representing the full combination of origin-destination pairs that are less than `max_dist` meters apart.
A common saying in data science is that 80% of the effort goes into the pre-processing stage.
This is equally true for spatial interaction modelling as it is for other types of data intensive analysis/modelling work.
So what does this function return?

```{r}
names(od)
```

As shown in the output above, the function allows you to use any variable in the origin or destination data in the function, with column names appended with `origin` and `destination`.

Note: support for OD datasets in which the origins and destinations are different objects is work in progress (see [#3](https://github.com/Robinlovelace/si/issues/3)).

## Feedback

We would be interested to hear how the approach presented in this package compared with other implementations such as those presented in the links below.
If anyone would like to try the approach or implement it in another language feel free to get in touch via the issue tracker.

## Further reading

For details on what SIMs are and how they have been defined mathematically and in code from first principles, see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims-first-principles.html).

To dive straight into using {si} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/si/articles/si.html).

For a detailed introduction to SIMs, support by reproducible R code, see Adam Dennett's [2018 paper](https://doi.org/10.37970/aps.v2i2.38).

## Other SIM packages

- The [`spflow` R package](https://github.com/LukeCe/spflow) on CRAN
- The [`spint` Python package](https://spint.readthedocs.io/en/latest/)
- The [`gravity`](https://cran.r-project.org/package=gravity) R package