### 2022-06-13 data-raw/si-classes.Rmd
Rename pkg with changer, close #20
Former-commit-id: 6c6fe93a3cbf6dac23e5eae172f3e6a4e4332fba

---
title: "Using the sim package with different input classes"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# remotes::install_github("robinlovelace/sim")
library(sim)
# ?si_predict
```

```{r}
od = si_to_od(si_zones, si_zones, max_dist = 4000)
m = lm(od$origin_all ~ od$origin_bicycle)
od_updated = si_predict(od, m)
class(od_updated)
od_dt = data.table::data.table(od)
class(od_dt)
od_updated_dt = si_predict(od_dt, m)
class(od_updated_dt)
identical(od_updated$interaction, od_updated_dt$interaction)
```