### 2022-06-13 data-raw/sim-paper.Rmd
Add draft paper content
Former-commit-id: aa491d3dc66d72527c90832a82313d39b48b6086

---
# title: "sim: A Package for Developing Spatial Interaction Models"
# output: rmarkdown::html_vignette
# vignette: >
#   %\VignetteIndexEntry{sim: A Package for Developing Spatial Interaction Models}
#   %\VignetteEngine{knitr::rmarkdown}
#   %\VignetteEncoding{UTF-8}

bibliography: paper.bib

output: rticles::jss_article
documentclass: jss
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: http://rstudio.com
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "\\pkg{sim}: A Package for Developing Spatial Interaction Models"
  # If you use tex in the formatted title, also supply version without
  plain:     "sim: A Package for Developing Spatial Interaction Models"
  # For running headers, if needed
  short:     "\\pkg{sim}: Spatial Interaction Models"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial interaction, movement, "\\proglang{R}"]
  plain:     [spatial interaction, movement, R]
preamble: >
  \usepackage{amsmath}
---

```{r echo=FALSE, eval=FALSE}
rbbt::bbt_update_bib(path_rmd = "vignettes/paper.Rmd", "vignettes/paper.bib")
```

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

# Introduction

Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
First developed by Alan Wilson in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
There are four main types of traditional SIMs [@wilson_family_1971]:

-   Unconstrained

-   Production-constrained

-   Attraction-constrained

-   Doubly-constrained

An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many iterations on this formulation):

$$
T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
$$ "where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.

Redefining the $W$ terms as $m$ and $n$ for origins and destinations respectively [@simini_universal_2012], this classic definition of the 'gravity model' can be written as follows:

$$
T_{i j}=K \frac{m_{i} n_{j}}{c_{i j}^{n}}
$$

For the purposes of this project, we will focus on production-constrained SIMs.
These can be defined as follows [@wilson_family_1971]:

$$
T_{ij} = A_iO_in_jf(c_{ij})
$$

where $A$ is a balancing factor defined as:

$$
A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
$$

$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.

More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
See @lenormand_systematic_2016 for a comparison of alternative approaches.

## Code formatting

In general, don't use Markdown, but use the more precise LaTeX commands instead:

-   \proglang{Java}

-   \pkg{plyr}

One exception is inline code, which can be written inside a pair of backticks (i.e., using the Markdown syntax).

If you want to use LaTeX commands in headers, you need to provide a `short-title` attribute.
You can also provide a custom identifier if necessary.
See the header of Section \ref{r-code} for example.

# \proglang{R} code {#r-code short-title="R code"}

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```

## Features specific to \pkg{rticles} {short-title="Features specific to rticles"}

-   Adding short titles to section headers is a feature specific to \pkg{rticles} (implemented via a Pandoc Lua filter). This feature is currently not supported by Pandoc and we will update this template if [it is officially supported in the future](https://github.com/jgm/pandoc/issues/4409).
-   Using the `\AND` syntax in the `author` field to add authors on a new line. This is a specific to the `rticles::jss_article` format.