### 2024-08-19 data-raw/schools_york.qmd
Merge code from Malcolm

# Modelling travel to school in York

The code below demonstrates how to fit a SIM with origins and destinations.
The data used in this example is from the city of York, UK, and is available in the `simodels` package.

Note: could become a vignette in the future.

```{r}
#| label: inputs-york
g1 = zones_york |>
  ggplot() +
  geom_sf() +
  labs(title = "Zones")
g2 = destinations_york |>
    ggplot() +
    geom_sf() +
    labs(title = "Destinations")
g1 + g2
```

Before we run any models let's compare the total number of pupils in the zones dataset and the destinations dataset (they should be the same):

```{r}
#| label: compare-totals
#| echo: true
zone_overestimate_factor = 
  (sum(zones_york$f0_to_15) + sum(zones_york$m0_to_15)) /
    sum(destinations_york$n_pupils)
zone_overestimate_factor
```

As one would expect, the total number of pupils in the zones dataset is a bit bigger than the total number of pupils in the destinations dataset: not all people aged 0 to 15 go to school, especially those under school age.
To tackle this issue we'll create a new variables called `pupils_estimated` in the zones dataset, which is the sum of the number of pupils in the zones dataset and the number of pupils in the destinations dataset.

```{r}
#| label: add-pupils-estimated
#| echo: true
zones_york = zones_york |>
  dplyr::mutate(
    pupils_estimated = (f0_to_15 + m0_to_15) / zone_overestimate_factor
  )
```

After the adjustment shown above, the totals in the origin and destination columns should be the same:

```{r}
#| label: compare-totals-after
#| echo: true
sum(zones_york$pupils_estimated)
sum(destinations_york$n_pupils)
```

Based on these inputs the `si_to_od()` function generates the OD data, as shown below (note: 2 versions are created, one with a maximum distance constraint for speed of processing, important when working with large datasets).

```{r}
#| echo: true
max_dist = 5000 # meters
od_from_si_full = simodels::si_to_od(zones_york, destinations_york)
od_from_si = simodels::si_to_od(zones_york, destinations_york, max_dist = max_dist)
```

```{r}
#| label: plot-od-all
#| layout-ncol: 2
m1 = od_from_si_full |>
  ggplot() +
  geom_sf(alpha = 0.1)
m2 = od_from_si |>
  ggplot() +
  geom_sf(alpha = 0.1)
m1
m2
```

The output OD dataset has column names taken from both the origin and destination datasets, with the following column names:

```{r}
names(od_from_si)
```

Let's run a simple model:

```{r}
#| label: simple-model-york
#| echo: true
gravity_model = function(beta, d, m, n) {
  m * n * exp(-beta * d / 1000)
} 
# perform SIM
od_res = simodels::si_calculate(
  od_from_si,
  fun = gravity_model,
  d = distance_euclidean,
  m = origin_pupils_estimated,
  n = destination_n_pupils,
#   constraint_production = origin_all,
  beta = 0.9
  )
```

We'll make one adjustment to the output dataset, renaming the `interaction` column to `trips`, and setting the total number of trips to be the same as the total number of pupils in the destinations dataset:

```{r}
#| label: adjust-output-york
#| echo: true
interaction_overestimate_factor = sum(destinations_york$n_pupils) / sum(od_res$interaction)
od_res = od_res |>
  dplyr::mutate(
    trips = interaction * interaction_overestimate_factor
  )
```

We can assess the model fit at three levels: the origin level (number of students departing from each zone), the destination level (the number arriving at each school in the input dataset) and the origin-destination level.

```{r}
#| label: r-squared-york
res_o = od_res |>
  group_by(O) |>
  summarise(
    Observed = first(origin_pupils_estimated),
    Modelled = sum(trips),
    Type = "Origin"
  )
res_d = od_res |>
  group_by(D) |>
  summarise(
    Observed = first(destination_n_pupils),
    Modelled = sum(trips),
    Type = "Destination"
  )
res_combined = bind_rows(res_o, res_d) |>
  # Create ordered factor with types:
  mutate(
    Type = factor(Type, levels = c("Origin", "Destination"))
  )
g_combined = res_combined |>
  ggplot() +
    geom_point(aes(x = Observed, y = Modelled)) +
    geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
    facet_wrap(~Type, scales = "free") +
    labs(
      title = "Model fit at origin, destination and OD levels (unconstrained)",
      x = "Observed",
      y = "Modelled"
    ) 
g_combined
```

Let's see if making the model production constrained can help:

```{r}
#| label: constrained-york
#| echo: true
od_res_constrained = simodels::si_calculate(
  od_from_si,
  fun = gravity_model,
  d = distance_euclidean,
  m = origin_pupils_estimated,
  n = destination_n_pupils,
  constraint_production = origin_pupils_estimated,
  beta = 0.9
  )
```

```{r}
#| label: r-squared-constrained
res_o = od_res_constrained |>
  group_by(O) |>
  summarise(
    Observed = first(origin_pupils_estimated),
    Modelled = sum(interaction),
    Type = "Origin"
  )
res_d = od_res_constrained |>
  group_by(D) |>
  summarise(
    Observed = first(destination_n_pupils),
    Modelled = sum(interaction),
    Type = "Destination"
  )
res_combined = bind_rows(res_o, res_d) |>
  mutate(
    Type = factor(Type, levels = c("Origin", "Destination"))
  )
res_combined |>
  ggplot() +
    geom_point(aes(x = Observed, y = Modelled)) +
    geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
    facet_wrap(~Type, scales = "free") +
    labs(
      title = "Model fit at origin, destination and OD levels (production constrained)",
      x = "Observed",
      y = "Modelled"
    )
```