### 2024-08-19 data-raw/si_oa_wpz.qmd

---
cache: true
---

```{r}
library(sf)
library(tidyverse)
library(tmap)
tmap_mode("view")
```

```{r}
od_data_oa_wpz_all = readr::read_csv("wf01aew_oa_v1.zip", col_names = c("OA11CD", "wz11cd", "n"))

# Workplace zone boundaries:
# https://geoportal.statistics.gov.uk/datasets/5ed6104f204944b6b60780842cbdaa5c_0/explore
wpz_all = sf::read_sf("Workplace_Zones_Dec_2011_GCB_in_England_and_Wales_2022_4617806297229452276.geojson")
wpz_all = wpz_all |>
  sf::st_transform("EPSG:4326")
names(wpz_all)
# [1] "OBJECTID" "wz11cd"   "lad11cd"  "lad11nm"  "lad11nmw" "GlobalID" "geometry"
# Keep only wz11cd and geometry:
wpz_all = wpz_all["wz11cd"]

summary(unique(od_data_oa_wpz_all$wz11cd) %in% wpz_all$wz11cd)
# and the other way around:
summary(unique(wpz_all$wz11cd) %in% od_data_oa_wpz_all$wz11cd)

# Take a 3km radius around Leeds:
sf::sf_use_s2(FALSE)
wpz_centroids = sf::st_centroid(wpz_all)
sf::sf_use_s2(TRUE)
leeds_boundary = zonebuilder::zb_zone("Leeds", n_circles = 2)
wpz_centroids_leeds = wpz_centroids[leeds_boundary, ]

# tmap::qtm(wpz_centroids_leeds)
si_wpz_leeds = wpz_all |>
  filter(wz11cd %in% wpz_centroids_leeds$wz11cd)
qtm(si_wpz_leeds)
```

```{r}
# Get OA data:
oa_all = sf::read_sf("/data/bronze/geoportal.gov.uk/Output_Areas_Dec_2011_Boundaries_EW_BGC_2022_7971430631129549274.gpkg")
oa_all = oa_all |>
  sf::st_transform("EPSG:4326")
oa_all_centroids = sf::st_centroid(oa_all)
oa_centroids_leeds = oa_all_centroids[leeds_boundary, ]
si_oa_leeds = oa_all |>
  filter(OA11CD %in% oa_centroids_leeds$OA11CD) |>
  select(OA11CD)
qtm(si_oa_leeds)
```

```{r}
#| label: od-data-leeds
si_oa_wpz = od_data_oa_wpz_all |>
  filter(wz11cd %in% si_wpz_leeds$wz11cd & OA11CD %in% si_oa_leeds$OA11CD)
sum(si_oa_wpz$n)
```

```{r}
#| label: od-origin-destination
si_oa_wpz_o = si_oa_wpz |>
  group_by(OA11CD) |>
  summarise(n = sum(n)) |>
  rename(n_o = n)
# join with origin data:
si_oa_wpz_o = si_oa_leeds |>
  left_join(si_oa_wpz_o) 
# tm_shape(si_oa_wpz_o) +
#   tm_polygons(col = "n_o")
plot(si_oa_wpz_o["n_o"])
```

```{r}
#| label: od-destination
si_oa_wpz_d = si_oa_wpz |>
  group_by(wz11cd) |>
  summarise(n = sum(n)) |>
  rename(n_d = n)

si_oa_wpz_d = si_wpz_leeds |>
    left_join(si_oa_wpz_d)

plot(si_oa_wpz_d["n_d"])
```

Now check the data:

```{r}
si_oa_wpz
si_oa_wpz_o
si_oa_wpz_d
```

More logical column names:

```{r}
si_oa_wpz_o = sf::st_sf(
  sf::st_drop_geometry(si_oa_wpz_o),
  geometry = sf::st_geometry(si_oa_wpz_o)
)
si_oa_wpz_d = sf::st_sf(
  sf::st_drop_geometry(si_oa_wpz_d),
  geometry = sf::st_geometry(si_oa_wpz_d)
)
sum(si_oa_wpz_o$n_o)
sum(si_oa_wpz_d$n_d)
sum(si_oa_wpz$n)
```

Save with usethis:

```{r}
usethis::use_data(si_oa_wpz, overwrite = TRUE)
usethis::use_data(si_oa_wpz_o, overwrite = TRUE)
usethis::use_data(si_oa_wpz_d, overwrite = TRUE)
```