### 2024-08-19 data-raw/test-si-to-od.qmd

---
format: gfm
---

```{bash}
#| eval: false
# Get data from the web
gh release 
```

```{r}
#| eval: false
piggyback::pb_list(repo = "itsleeds/atumie")
piggyback::pb_download("zones_test.gpkg", repo = "itsleeds/atumie")
piggyback::pb_download("social_grid.gpkg", repo = "itsleeds/atumie")
piggyback::pb_upload("zones_test.gpkg")
piggyback::pb_upload("social_grid.gpkg")
# Get urls from v0.0.5 release:
url_zones = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/zones_test.gpkg"
f_zones = basename(url_zones)
download.file(url_zones, f_zones)
url_grid = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/social_grid.gpkg"
download.file(url_grid, basename(url_grid))
```

```{r}
url_zones = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/zones_test.gpkg"
zones = sf::read_sf(url_zones)
url_grid = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/social_grid.gpkg"
grid = sf::read_sf(url_grid)
```

```{r}
devtools::load_all()
od = si_to_od(zones, grid)
```

```{r}
zonesg = sf::st_sf(
  sf::st_drop_geometry(zones),
  geometry = zones$geom
)
gridg = sf::st_sf(
  sf::st_drop_geometry(grid),
  geometry = grid$geom
)
odg = si_to_od(zonesg, gridg)

gridd = gridg[!duplicated(gridg$grid_id), ]

odg = si_to_od(zonesg, gridd)
```

```{r}
od = od::points_to_od(zones, grid)
```